---
title: "Data Visualization Examples"
author: "Erica Bishop"
date: "`r Sys.Date()`"
output: html_document
---
# Data Visualization Examples

Below are examples of visualizations taken from my various coursework. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(spData)
library(leaflegend)
library(leaflet)
library(tidyverse)
library(patchwork)

```

## Investigating spotify data with ggplot

These ggplots are from the same spotify data wrangled in bishop_data_wrangling.RMD. The goal of the visualization below was to quickly compare the difference in means across multiple features between two three user classifications - songs in my saved track lists on spotify, songs in a classmate's saved track list, and songs we had in common.

```{r echo=FALSE}
#read in data
eb_ja_tracks <- read_csv(here::here("eb_ja_tracks.csv"))

```

```{r}

#compare danceability
dance_plot <- ggplot(data = eb_ja_tracks,
       aes(y = danceability,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Danceability") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank()
  )

#compare energy
energy_plot <- ggplot(data = eb_ja_tracks,
       aes(y = energy,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Energy") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank()
  )

#compare key
key_plot <- ggplot(data = eb_ja_tracks,
       aes(y = key,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Key") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank()
  )

#compare valence
valence_plot <- ggplot(data = eb_ja_tracks,
       aes(y = valence,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Valence") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank()
  )

#compare tempo
tempo_plot <- ggplot(data = eb_ja_tracks,
       aes(y = tempo,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Tempo") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank()
  )

#compare mode
mode_plot <- ggplot(data = eb_ja_tracks,
       aes(y = loudness,
           x = user,
           col = user)) +
  geom_boxplot() +
  labs(title = "Loudness") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank()
  )


var_compare_plot <- dance_plot + energy_plot + key_plot + valence_plot + tempo_plot + mode_plot +
  plot_layout(guides = 'collect')

var_compare_plot

```

## Spatial data visualization with `leaflet`

The map below shows the world population by country using the `world` data from the `spData` package.

```{r}

pal = colorNumeric(palette = "PuRd", domain = world$pop) #create a color palette

leaflet(data = world) |> #add data
  addPolygons(data = world,
              color = ~pal(pop), #color by population
              fillOpacity = 0.5) |>
  addProviderTiles(providers$OpenStreetMap) |> #add known basemap
  addPolygons(weight = 0.5,
              color = "black") |> 
  addLegendNumeric(pal = pal, #create more intuitive legend
            values = ~pop,
            position = 'topright',
            title = "Population",
            decreasing = TRUE
            ) |> 
  setView(lng = -120, lat = 47, zoom = 3) #toggle where map view starts and zoom
  

```
